---
title: "cross-validation"
author: "cathy"
date: "2024-11-12"
output: html_document
---
```{r}
library(tidyverse)
library(modelr)
library(mgcv)

#install.packages(SemiPar)
library(SemiPar)
set.seed(1)
```

look at LIDAR data

```{r}
data("lidar")

lidar_df =
  lidar |> 
  as_tibble() |> 
  mutate(id = row_number())
```

```{r}
lidar_df |> 
  ggplot(aes(x = range, y = logratio)) +
  geom_point()
```

##Try to do CV

we'll compare 3 mdoels  -- one linear, one smooth, one wiggly

constrct training and testing df
```{r}
train_df = sample_frac(lidar_df, size = .8)
test_df = anti_join(lidar_df, train_df, by = "id") # anything in the lidar_df but not in train_df
```

look at these
```{r}
ggplot(train_df, aes(x = range, y = logratio)) +
  geom_point() +
  geom_point(data = test_df, color = "red")
```


fit 3 models
```{r}
linear_mod = lm(logratio ~ range, data = train_df)
smooth_mod = gam(logratio ~ s(range), data = train_df)
wiggly_mod = gam(logratio ~ s(range, k = 30), sp = 10e-6, data = train_df)
```

look at the fit
```{r}
train_df |> 
  add_predictions(smooth_mod) |> 
  ggplot(aes(x = range, y = logratio)) +
  geom_point() +
 # geom_point(data = test_df, color = "red") +
  geom_line(aes(y = pred), color = "red") 
```

compare these numerically using RMSE
```{r}
rmse(linear_mod, test_df)
rmse(smooth_mod, test_df)
rmse(wiggly_mod, test_df)
```


## repeat the train / test

```{r}
cv_df = 
  crossv_mc(lidar_df, 100) |> 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )
```

```{r}
cv_df |> 
  pull(train) |> 
  nth(3) |> 
  as_tibble()
```

fit models, extract RMSEs

```{r}
cv_res_df = 
  cv_df |> 
  mutate(
    linear_mod = map (train, \(x)lm(logratio ~ range, data = x)),
    smooth_mod = map (train, \(x)lm(logratio ~ s(range), data = x)),
    wiggly_mod = map (train, \(x)lm(logratio ~ s(range), k = 30, sp = 10e-6, data = x))) |> 
  mutate(
    rmse_linear = map2_dbl(linear_mod, test, rmse),
    rmse_smooth = map2_dbl(smooth_mod, test, rmse),
    rmse_wiggly = map2_dbl(wiggly_mod, test, rmse)
    
  )

```


look at RSMD contirbution
```{r}
cv_res_df |> 
  select(starts_with(rmse)) |> 
  pivot_longer(
    everything(),
    names_to = "model",
    names_prefix = "rmse_"
  ) |> 
  ggpolot(aes(x =  model, y = rmse)) +
  geom_violin()
```



## nepalese, children df
```{r}
child_df = 
  read.csv("data/nepalese_children.csv") |> 
  mutate(
    weight_ch7 = (weight > 7) * (weight -7)
  )
  
```

look at data
```{r}
child_df |> 
  ggplot(aes(x = weight, y = armc)) +
  geom_point(alpha = .5)
```

fit some models
```{r}
linear_mod = lm(armc ~ weight, data = child_df)
pwl_mod = lm(armc ~ weight + weight_ch7, data = child_df)
smooth_mod = gam(armc ~ s(weight), data = child_df)
```


look at the models
```{r}
child_df |> 
  add_predictions(pwl_mod) |> 
  ggplot(aes(x = weight, y = armc)) +
  geom_point(alpha = .5)+
  geom_line(aes(y = pred), color = "red")
```

CV to slecet models
```{r}
cv_df. =
crossv_mc(child_df, 100) |> 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )
```

apply models and extract RMSE

```{r}
cv_res_df = 
  cv_df |> 
  mutate(
    linear_mod = map(train, \(x) lm (armc ~ weight, data = x)),
    psl_mod = map(train, \(x) lm (armc ~ weight + weight_ch7, data = x)),
    smooth_mod = map(train, \(x) lm (armc ~ weight + weight_ch7, data = x))
  ) |> 
  mutate(
    rmse_linear = map2_dbl(linear_mod, test, rmse),
    rmse_psl = map2_dbl(pwl_mod, test, rmse),
    rmse_linear = map2_dbl(linear_mod, test, rmse),
  )
```

