---
title: "Untitled"
author: "cathy"
date: "2024-11-07"
output: github_document
---

```{r}
library(tidyverse)
library(p8105.datasets)

set.seed(1)

```

do some data cleaning

```{r}
data("nyc_airbnb")

nyc_airbnb = 
  nyc_airbnb |> 
  mutate(stars = review_scores_location / 2) |> 
  rename(
    borough = neighbourhood_group,
    neighborhood = neighbourhood) |> 
  filter(borough != "Staten Island") |> 
  select(price, stars, borough, neighborhood, room_type) |> 
  mutate(
    borough = fct_infreq(borough),
    room_type = fct_infreq(room_type)
  )
```

## fit some models

```{r}
fit = lm(price ~ stars, data = nyc_airbnb)

summary(fit)

names(summary(fit))

coef(fit)

fit |> 
  broom::tidy() |>  # format as a table
  select(term, estimate, p.value)

fit |> 
  broom::glance()
```

try a bit more complex
```{r}
fit = 
  lm(price ~ stars + borough, data = nyc_airbnb) 

fit |> 
  broom::tidy() |> 
  select(term, estimate, p.value) |> 
  mutate(
    term =  str_replace(term, "borough", "Borough:")
  ) |> 
  knitr::kable(digits = 3)
```

## some diagnositics

most diagnostics use residuals.
```{r}
modelr::add_residuals(nyc_airbnb, fit) |> 
  ggplot(aes(x = resid))+
  geom_histogram()
```


```{r}
modelr::add_residuals(nyc_airbnb, fit) |> 
  ggplot(aes(x = borough, y = resid)) +
  geom_violin() +
  ylim(-100, 100)
```


```{r}
modelr::add_residuals(nyc_airbnb, fit) |> 
  ggplot(aes(x = stars, y = resid)) +
  geom_point()
```


residuals against fitted values
```{r}
nyc_airbnb |> 
  modelr::add_residuals(fit) |> 
  modelr::add_predictions(fit) |> 
  ggplot(aes(x = pred, y = resid)) +
  geom_point()
```


## Hypothiesis testing

```{r}
fit |> 
  broom::tidy()

fit_null = lm(price ~ stars + borough, data = nyc_airbnb)
fit_alt = lm(price ~ stars + borough + room_type, data = nyc_airbnb)
```

look at both
```{r}
fit_null |> 
  broom::tidy()

fit_alt |> 
  broom::tidy()

anova(fit_null, fit_alt) |> 
  broom::tidy()
```

## Do effects differ across boroughs

First, use a lot of interactions
```{r}
nyc_airbnb |> 
  lm(price ~ stars * borough + room_type * borough, data = _) |> 
  broom::tidy()
```

could just fit separate models
```{r}
nyc_airbnb |> 
  filter(borough == "Manhattan") |> 
  lm(price ~ stars + room_type, data = _) |> 
  broom::tidy()

nyc_airbnb |> 
  filter(borough == "Brooklyn") |> 
  lm(price ~ stars + room_type, data = _) |> 
  broom::tidy()
```

get fancy and use list colunms
```{r}
nyc_airbnb |> 
  nest(data = -borough) |> 
  mutate(
    model = map(data, \(x) lm(price ~ stars + room_type, data = x)),
    results = map(model, broom::tidy)
  ) |> 
  select(borough, results) |> 
  unnest(results) |> 
  select(borough, term, estimate) |> 
  pivot_wider(
    names_from = term,
    values_form = estimate
  )
```


what about room tyoe across manhattan neighborhoods?
```{r}
nyc_airbnb |> 
  filter(
    borough == "Manhattan",
    neighborhood == "Chinatown") |> 
  lm(price ~ stars + room_type, data = _) |> 
  broom::tidy()

nyc_airbnb |> 
  filter(
    borough == "Manhattan",
    neighborhood == "Chelsea") |> 
  lm(price ~ stars + room_type, data = _) |> 
  broom::tidy()

```  
  
Let's do this for all neighborhooods!

```{r}
manhattan_neighborhood_fit_df =
nyc_airbnb |> 
  filter(borough == "Manhattan") |> 
  nest(data = -(borough:neighborhood)) |> 
  mutate(
    model = map(data, \(x) lm(price ~ stars + room_type, data = x)),
    resluts = map(model, broom::tidy)
  ) |> 
  select(neighborhood, results) |> 
  unnest(results)
```

  
look at the effect of room type
```{r}
manhattan_neighborhood_fit_df |> 
  filter(str_detect(term, "room_type")) |> 
  ggplot(aes(x = term, y = esitimate))+
  geom_boxpot()
```


```{r}
manhattan_neighborhood_fit_df |> 
  filter(str_detect(term, "room_type")) |> 
  filter(estimat > 0)

nyc_airbnb |> 
  filter(
    neighborhood == "NoHo",
    room_type == "Shared room"
  )
```

